Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties

### 2022-02-24 papers/paper/preprint.tex
better figures color schemes and some updates to text

\begin{abstract}
    \textbf{Draft in preparation to Earth's Future}.
\end{abstract}

@@

\begin{table}
    \centering
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \ref{sec:ead}         \\
        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 & TBD                   \\
        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                & TBD                   \\
        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   & TBD                   \\
        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       & TBD                   \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=4in]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a single decision (yellow).
        For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
    }\label{fig:xlrm}
\end{figure}

@@

\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    %\includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:msl-boxplots}
\end{figure}

For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.

@@

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
This time series is shown in \cref{fig:observed-surges}(a).
In (b) we use the Weibull plotting position
\begin{equation}\label{eq:plot-pos}
    \frac{r}{N + 1}
\end{equation}
where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).

\begin{figure}
    \centering
    %\includegraphics[height=1.5in]{historic_surge.pdf}%
    %\includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
        (a):
        Time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

@@

There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
\begin{equation}
    \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
\end{equation}
where $y_t = \overline{y}_t + y'_t$ is the annual maximum flood, relative to the gauge, in year $t$, and $D(h - y_t)$ is the damage as a function of depth relative to the house.
Following \citet{zarekarizi_suboptimal:2020} we use the HAZUS deterministic depth-damage relationship to parameterize $D(h-y_t)$ (see fig. XYZ(c)).\james{Cite HAZUS}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
    \caption{
        Outline of the proposed decision theoretic framework.
        Boxes are labeled in the order in which they are presented in the text.
        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
        In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
        Finally in \cref{sec:synthesizing} we add boxes (f) and (g) to synthesize across multiple \glspl{pdf}.
    }\label{fig:flowchart}
\end{figure}

@@

\begin{figure}
    %\includegraphics[width=\textwidth]{scenario_maps.png}
    \caption{
        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
        One dot represents one \gls{sow}.
        %TODO
    }\label{fig:scenario-map}
\end{figure}

@@

An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.

@@

\begin{figure}
    \centering
    %\includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    %\includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Need a summary sentence.
        %TODO
        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}

For example, \cref{fig:tradeoffs}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
Similarly, \cref{fig:tradeoffs} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp} 8.6 with fast ice sheet dynamics).
Consistent with \cref{fig:scenario-map}, the variability of lifetime costs increases as $\Delta h$ increases.
Once $\Delta h$ reaches \SIrange[]{3}{7}{ft}, depending on the model considered, construction costs start to dominate flood losses, and thus higher values of $\Delta h$ increase average lifetime costs.

Calculating either of the quantities shown in \cref{fig:tradeoffs} \emph{requires synthesizing across \glspl{sow}}.
However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
This means that any assessment of decision performance (\eg, an optimal strategy, a Pareto frontier, or a cost-benefit analysis) is conditional upon an (explicit or implicit) model of future outcomes.
Where these models are implicit, they should be made more explicit to facilitate iterative model critique and improvement.

@@

We apply a grid-based discretization scheme to calculate these weights.
First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
As before, we take this space to be the mean sea level in 2100.
Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:weighting}.
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
    \end{cases}
\end{equation}
This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{../img/weighting.pdf}
    \caption{
        Schematic of re-weighting scheme defined in \cref{eq:weights} for $J=4$.
        Top panel shows $p_\mathrm{true}$ and bottom panel shows $F_\mathrm{true}$.
        Vertical black lines show $\frac{1}{2}\qty[\psi_j + \psi_{j+1}]$ for $j=1, \ldots, J-1$.
        Green text shows $w_1, \ldots, w_4$ and how they are calculated from $F_\mathrm{true}$.
        Need to update this figure using real data.
        %TODO
    }\label{fig:weighting}
\end{figure}

Need to explain how we came up with our two or three subjective priors.\klaus{Would love suggestions here!}
Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
\Cref{fig:tradeoffs} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')

\begin{figure}
    \centering
    %\includegraphics[width=3in]{lsl_priors.pdf}~
    %\includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
    \caption{
        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
        Alternative priors can provide robustness checks.
        Need to add panel labels, update data, \etc.
        %TODO
    }\label{fig:tradeoffs}
\end{figure}

This technique can also be used as a diagnostic tool.
In \cref{fig:model-average-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}

\begin{figure}
    \centering
    %\includegraphics[width=6in]{inference_weights.pdf}
    \caption{
        The average weight assigned to \glspl{sow} from each model (\gls{rcp} scenario and ice sheet parameterization) for both priors considered.
        If the sampling distribution (green line) is used, then the weight assigned to each \gls{sow} is the same, regardless of whether it is over- or under-sampled.
    }\label{fig:model-average-weights}
\end{figure}

@@

Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
We introduce an approach to bridge this divide through a computationally efficient method that weights each state of the world based on a probability distribution over possible futures.
Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.

@@

\begin{enumerate}
    \item Supplementary figures and analysis online
    \item Link to live repository on GitHub
    \item Link to code DOI on Zenodo
\end{enumerate}

@@

\begin{enumerate}
    \item MARISA\james{details}
    \item PSIRC\james{details}
    \item Rice University (for startup)
    \item Dartmouth College (for startup)
    \item Tor Erlend Fjelde for helpful advice on implementing the \gls{gev} prior using \texttt{Turing.jl}.
    \item For many journals, a code and data statement will go here.\james{Keep an eye on this}
\end{enumerate}

@@

See journal requirements and format\james{Keep an eye on this}

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-expected-damage-emulator}
    \caption{
        As discussed in XXX, we model expected annual damages as a function of the house's elevation relative to \gls{lsl}. % TODO: ref equation
    }\label{fig:cost-expected-damage-emulator}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-up-front}
    \caption{
        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial fee.
        Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
    }\label{fig:cost-up-front}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{grid-sketch}
    \caption{
        Schematic of equation XXX used to calculate weights for each sample. % TODO: reference equation
        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
        As discussed in section XXX, the weights (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution  (horizontal dotted lines) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines). % TODO: reference section
    }\label{fig:grid-sketch}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
    }\label{fig:lsl-evolution}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-priors-weights}
    \caption{
        Subjective priors for local sea level.
        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
        The \glspl{pdf} of these subjective priors are shown in panel (A).
        In panels (B)-(D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
        Specifically, (B)-(D) show the average weight given to each model by each of the three subjective priors.
        The ``Slow SLR'' prior places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
        The ``Fast SLR'' prior places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
        The ``Intermediate SLR'' scenario places approximately equal weight on all models.
    }\label{fig:lsl-priors-weights}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{scenario-map-height-slr}
    \caption{
        Expected total lifetime cost (damages plus up-front cost; eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
        The best outcomes (yellow) arise when the house is not elevated and sea level rise is minimal (bottom left corner).
        The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
        In general, a height increase of $<\SI{3}{ft}$ can be considered unrealistic \citep{zarekarizi_suboptimal:2020}.
        Values are sensitive to model constants; see \cref{tab:uncertainties}.
    }\label{fig:scenario-map-height-slr}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
    \caption{
        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost; eq.XXX) as a function of \gls{slr} in 2100 for several values of $\Delta h$ and $h_0$.
        Colors indicate the density of \glspl{sow}; this data is often presented as a scatterplot (CITE) but we bin to a 2D histogram. % TODO: cite scenario maps
        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
        In all cases, elevating the house reduces the variance in total lifetime cost.
        Values are sensitive to model constants; see \cref{tab:uncertainties}.
    }\label{fig:scenario-map-slr-cost}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{surge-gev-priors}
    \caption{
        Prior distributions for annual maximum storm surge.
        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following REFERENCE. % TODO: cite
        See ex.~XYQ for details. % TODO: cross-ref equation
        We apply priors over the 2, 10, 100, and 500 year events.
    }\label{fig:surge-gev-priors}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE). % TODO: cite source
        (A):
        Time series of historic storms.
        Blue (green) arrows denote notable tropical cyclones (nor'easters).
        (B):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian GEV fit.
    }\label{fig:surge-obs-return}
\end{figure}

>>>>

\begin{abstract}
    The policy and engineering tools used to inform climate adaptation etc rely on standards and cost-benefit analyses that depend on risk probabilities.
    One example is house elevation: the recommended elevation is a standard (BFE plus a foot) and funding decisions rely on CBA.
    Under stationarity, the probabilities for this can almost be agreed upon.
    Under nonstationarity, there are deep uncertainties.
    We show that the way these deep uncertainties are synthesized into CBA matters a lot.
    Thus, static decisions are subject to deep uncertainties
\end{abstract}

@@

\begin{table}
    \centering
    \caption{
        Summary of parameters, their notation, and how their uncertainty is modeled.
    }\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name             & Symbol           & Uncertainty                                                                          & See more                     \\
        \midrule
        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \cref{sec:sea-level}         \\
        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \cref{sec:storm-surge}       \\
        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \cref{sec:led}               \\
        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       &                              \\
        \bottomrule
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=4in]{xlrm.pdf}
    \caption{
        Conceptual diagram of the decision problem.
        A \gls{sow} consists of a description of the uncertain factors (red).
        We model a single decision (yellow).
        For each combination of \gls{sow} and policy, the system model (blue) is used to calculate metrics (orange) describing the performance of the policy over the selected \gls{sow}.
    }\label{fig:xlrm}
\end{figure}

@@

\Cref{fig:lsl-evolution} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-evolution}
    \caption{
        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
    }\label{fig:lsl-evolution}
\end{figure}

For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \gls{sow}}.

@@

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
This time series is shown in \cref{fig:surge-obs-return}(a).
In (b) we use the Weibull plotting position
\begin{equation}\label{eq:plot-pos}
    \frac{r}{N + 1}
\end{equation}
where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{surge-obs-return}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (A):
        Time series of historic storms.
        Blue (green) arrows denote notable tropical cyclones (nor'easters).
        (B):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit.
    }\label{fig:surge-obs-return}
\end{figure}

@@

There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
\begin{equation}\label{eq:ead}
    \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
\end{equation}
where $y_t = \overline{y}_t + y'_t$ is the annual maximum flood, relative to the gauge, in year $t$, and $D(h - y_t)$ is the damage as a function of depth relative to the house.
Following \citet{zarekarizi_suboptimal:2020} we use the HAZUS deterministic depth-damage relationship to parameterize $D(h-y_t)$ (see fig. XYZ(c)).\james{Cite HAZUS}

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{bayes-rdm.pdf}
    \caption{
        Outline of the proposed decision theoretic framework.
        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
        Next in \cref{sec:multiple-pdf} we use scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem'' (region 2).
        Finally in \cref{sec:synthesizing} propose a method for weighting \glspl{sow} to synthesize across multiple \glspl{pdf} (region 3).
    }\label{fig:flowchart}
\end{figure}

@@

@@

An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:lsl-evolution}.
This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
    \caption{
        Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
        (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
        (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
        Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
        Colored lines highlight three representative models for emphasis.
    }\label{fig:tradeoffs-by-rcp}
\end{figure}

For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
Similarly, \cref{fig:tradeoffs-by-rcp} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp} 8.6 with fast ice sheet dynamics).
Consistent with \cref{fig:scenario-map}, the variability of lifetime costs increases as $\Delta h$ increases.
Once $\Delta h$ reaches \SIrange[]{3}{7}{ft}, depending on the model considered, construction costs start to dominate flood losses, and thus higher values of $\Delta h$ increase average lifetime costs.

Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} \emph{requires synthesizing across \glspl{sow}}.
However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
This means that any assessment of decision performance (\eg, an optimal strategy, a Pareto frontier, or a cost-benefit analysis) is conditional upon an (explicit or implicit) model of future outcomes.
Where these models are implicit, they should be made more explicit to facilitate iterative model critique and improvement.

@@

We apply a grid-based discretization scheme to calculate these weights.
First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
As before, we take this space to be the mean sea level in 2100.
Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:grid-sketch}.
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
    \end{cases}
\end{equation}
This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{grid-sketch}
    \caption{
        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
    }\label{fig:grid-sketch}
\end{figure}

Need to explain how we came up with our two or three subjective priors.\klaus{Would love suggestions here!}
Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
\Cref{fig:tradeoffs-by-prior} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')

This technique can also be used as a diagnostic tool.
In \cref{fig:lsl-priors-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{lsl-priors-weights}
    \caption{
        Subjective priors for local sea level.
        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
        The \glspl{pdf} of these subjective priors are shown in panel (A).
        In panels (B-D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
        Specifically, (B-D) show the average weight given to each model by each of the three subjective priors.
        The ``Slow SLR'' prior (B) places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
        The ``Intermediate SLR'' (C) prior places approximately equal weight on all models.
        The ``Fast SLR'' prior (D) places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
    }\label{fig:lsl-priors-weights}
\end{figure}

@@

Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
We introduce an approach to bridge this divide through a computationally efficient method that weights each \gls{sow} based on a probability distribution over possible futures.
Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.

@@

@@

This work was supported by the National Oceanic and Atmospheric Administration (NOAA) through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
JDG thanks Rice University for support.
KK thanks Dartmouth College for support.
The authors thank Tor Erlend Fjelde for helpful comments.

\subsection*{Code and data availability}

All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
This code is written in the open source Julia programming language.

@@

JDG and KK designed the research.
JDG wrote the codes and ran simulations.
JDG wrote the manuscript with support from KK.
JDG wrote the manuscript in consulation and collaboration with KK.

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-expected-damage-emulator}
    \caption{
        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{lsl}.
        Damages ($y$ axis) are shown as a percentage of house value.
    }\label{fig:cost-expected-damage-emulator}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{cost-up-front}
    \caption{
        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial cost.
        Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
    }\label{fig:cost-up-front}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=4in]{surge-gev-priors}
    \caption{
        Prior distributions for annual maximum storm surge.
        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
        See \cref{sec:storm-surge} for details.
        For the 2, 10, 100, and 500 year events we apply Normal distributions, truncated at zero, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
    }\label{fig:surge-gev-priors}
\end{figure}
